---
name: Abhishek
surname: Kumar
position: "Senior Research Fellow"
address: "Department of Botany, Panjab University, Chandigarh"
profilepic: square.jpg
#phone: +91 90344 73233
email: "abhikumar.pu@gmail.com"
www: akumar.netlify.app
#orcid: 0000-0003-2252-7623
#researchgate: Abhishek-Kumar-331
twitter: abkumar_
github: kumar-a
#linkedin: abhikumar86
date: "`r format(Sys.time(), '%B %Y')`"
headcolor: "414141"

header-includes:
  - \definecolor{linkscolor}{HTML}{026dbb}
  - \hypersetup{colorlinks=true, linkcolor=linkscolor, filecolor=linkscolor, urlcolor=linkscolor, urlbordercolor=linkscolor}

output: 
  vitae::awesomecv:
    page_total: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
library(dplyr)
library(fontawesome)
library(tibble)
# remotes::install_github("mitchelloharawild/vitae")
library(vitae)
```


# Education

```{r}

tribble(
  ~Degree, ~Year, ~ Institution,
  "Doctor of Philosophy (Botany/Plant Ecology)", "2017 -- 2024", "Panjab University, Chandigarh, IN",
  "Master of Science (Botany)", "2015 -- 2017", "Panjab University, Chandigarh, IN",
  "Bachelor of Science (Botany, Chemistry, Zoology)", "2012 -- 2015", "Arya PG College, Panipat, IN"
) |> 
  detailed_entries(what = Institution, where = Year, with = Degree)
```

# Research Interests

- **Macroecology** Patterns and determinants of plant species richness along elevational gradients

- **Population Ecology** Climate-related redistribution of mountain trees (*Pinus*, *Quercus*, and *Rhododendron*)

- **Ecosystem Ecology** Patterns and determinants of litter decomposition in terrestrial ecosystems

- **Restoration Ecology** Ecological restoration of degraded ecosystems and coal mine spoils

- **Research Synthesis** Bibliometrics, systematic reviews, and meta-analysis

# Publications

A complete list of publications is available from my ORCID (<https://orcid.org/0000-0003-2252-7623>)

```{r}
# bibliography_entries("mypubs.bib") |> 
#     filter(type == "article-journal") |>
#     arrange(desc(issued))
```

## Journal articles

- **Kumar, A.**, Patil, M., Kumar, P., & Singh, A. N. (2024b). Exploring elevational patterns of plant species richness: Insights from Western Himalayas. *Ecology and Evolution* (accepted)

- **Kumar, A.**, Patil, M., Kumar, P., & Singh, A. N. (2024a). Determinants of plant species richness along elevational gradients: Insights with climate, energy and water-energy dynamics. *Ecological Processes 13*, 86. <https://doi.org/10.1186/s13717-024-00563-z>

- **Kumar, A.**, Kumar, P., Patil, M., Hussain, S., Yadav, R., Sharma, S., Tokas, D., Singh, S., & Singh, A. N. (2024). Disturbance and vegetational structure in an urban forest of Indian Siwaliks: an ecological assessment. *Environmental Monitoring and Assessment, 196*(8), 691. <https://doi.org/10.1007/s10661-024-12801-0>

- **Kumar, A.**, Patil, M., Kumar, P., Kumar, M., & Singh, A. N. (2022). Plant ecology in Indian Siwalik range: A systematic map and its bibliometric analysis. *Tropical Ecology, 63*(3), 338--350. <https://doi.org/10.1007/s42965-022-00229-x>

- Singh, A. N., & **Kumar, A.** (2022b). Comparative soil restoration potential of exotic and native woody plantations on coal mine spoil in a dry tropical environment of India: A case study. *Land Degradation & Development, 33*(12), 1971--1984. <https://doi.org/10.1002/ldr.4286>

- Singh, A. N., & **Kumar, A.** (2022a). Ecological performances of exotic and native woody species on coal mine spoil in Indian dry tropical region. *Ecological Engineering, 174*, 106470. <https://doi.org/10.1016/j.ecoleng.2021.106470>

- Patil, M., **Kumar, A.**, Kumar, P., Cheema, N. K., Kaur, R., Bhatti, R., & Singh, A. N. (2020). Comparative litter decomposability traits of selected native and exotic woody species from an urban environment of north-western Siwalik region, India. *Scientific Reports, 10*, 7888. <https://doi.org/10.1038/s41598-020-64576-2>

## Book chapters

- **Kumar, A.**, Patil, M., Kumar, P., & Singh, A. N. (2021). Phosphorus and litter decomposability traits in tropical forest ecosystems under changing environment: A synthesis. In R. K. Chaturvedi, R. Singh, & R. Bhadouria (Eds.), *Tropical Dry Forests: Emerging Features and Ecological Perspectives* (pp. 311–336). Nova Science Publishers, New York.

- **Kumar, A.**, Yadav, R., Patil, M., Kumar, P., Zhang, L., Kaur, A., Sharma, S., Hussain, S., Tokas, D., & Singh, A. N. (2020). Sustainable management of national parks and protected areas for conserving biodiversity in India. In L. Zhang (Ed.), *Advances in Forest Management under Global Change* (pp. 75–91). IntechOpen, London. <https://doi.org/10.5772/intechopen.92435>

# Skills

A full list of accomplishments and certificates is available from my webpage (<https://akumar.netlify.app>)

- `r fontawesome::fa(name = "r-project", fill = "#026dbb")` statistical environment and `RStudio`
- Data analysis and visualization (`ggplot2`, `MuMIn`, `stats`, `tidyverse`)
- Linear and non-linear mixed-effects modelling (`nlme`)
- Spatial data analysis (`bfast`, `MODISTools`, `sf`, `terra`, `tmap`)
- Species distribution modelling (`biomod2`, `MaxEnt`)
- Species diversity and multivariate analysis (`ade4`, `iNEXT`, `vegan`)
- Structural equation modelling (`piecewiseSEM`)
- Systematic reviews and meta-analysis (`bibliometrix`, `metafor`, `revtools`)
- Reproducible research (`git`, `quarto`, `rmarkdown`, `knitr`)


# Awards and Distinctions

```{r dist}
# what, when, with, where, why
tribble(
  ~what, ~date, ~org, 
  "Young Scientist Award", "2022", "Punjab Academy of Sciences, Patiala",
  "Best Poster Award", "2020", "National Conference on Climate Change, CSIR-NBRI, Lucknow",
  "Junior Research Fellowship", "2017", "University Grants Commission, New Delhi") |> 
    detailed_entries(what = org, when = date, with = what, .protect = FALSE)
```

# Professional Memberships

```{r}
tribble(
  ~Membership, ~Year, ~Institution, ~Where,
  "Student member (#521)", "2023 -- Present", "Society of Open, Reliable and Transparent Ecology and Evolutionary Biology (SORTEE)", "Salem, Oregon, US",
  "Life member (#L-1599)", "2022 -- Present", "Punjab Academy of Sciences (PAS)", "Patiala, Punjab, IN",
  "Student member (#67527876)", "2021 -- 2025", "International Association for Vegetation Science (IAVS)", "Bethesda, Maryland, US"
) |> 
  detailed_entries(Membership, Year, Institution, Where)
```

# Academic Services

Peer-reviewed for (i) *Ecological Engineering*; (ii) *Ecology and Evolution*; (iii) *Ecosphere*; (iv) *Frontiers in Forests and Global Change*; (v) *Oikos*; (vi) *One Ecosystem* and other reputed journals
