---
name: Abhishek
surname: Kumar
position: "Senior Research Fellow"
address: "Department of Botany, Panjab University, Chandigarh"
profilepic: square.jpg
#phone: +91 90344 73233
email: "abhikumar.pu@gmail.com"
www: akumar.netlify.app
#orcid: 0000-0003-2252-7623
#researchgate: Abhishek-Kumar-331
twitter: abkumar_
github: kumar-a
#linkedin: abhikumar86
date: "`r format(Sys.time(), '%B %Y')`"
headcolor: "414141"

header-includes:
  - \definecolor{linkscolor}{HTML}{026dbb}
  - \hypersetup{colorlinks=true, linkcolor=linkscolor, filecolor=linkscolor, urlcolor=linkscolor, urlbordercolor=linkscolor}

output: 
  vitae::awesomecv:
    page_total: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
library(dplyr)
library(tibble)

# remotes::install_github("mitchelloharawild/vitae")

library(vitae)
```


# Education

```{r}

tribble(
  ~Degree, ~Year, ~ Institution,
  "Doctor of Philosophy (Botany/Plant Ecology)", "2017-Present", "Panjab University, Chandigarh, IN",
  "Master of Science (Botany)", "2015-2017", "Panjab University, Chandigarh, IN",
  "Bachelor of Science (Botany, Chemistry, Zoology)", "2012-2015", "Arya PG College, Panipat, IN"
) |> 
  detailed_entries(what = Institution, where = Year, with = Degree)
```

# Awards and Distinctions

```{r dist}
# what, when, with, where, why
tribble(
  ~what, ~date, ~org, 
  "Young Scientist Award", "2022", "Punjab Academy of Sciences, Patiala",
  "Best Poster Award", "2020", "National Conference on Climate Change, CSIR-NBRI, Lucknow",
  "Junior Research Fellowship", "2017", "University Grants Commission, New Delhi") |> 
    detailed_entries(what = org, when = date, with = what, .protect = FALSE)
```

# Research Interests

- **Macroecology** Elevational patterns and determinants of plant species richness

- **Population Ecology** Climate-related redistribution of mountain trees (*Pinus*, *Quercus* and *Rhododendron*)

- **Ecosystem Ecology** Patterns and determinants of litter decomposition in terrestrial ecosystems

- **Restoration Ecology** Ecological restoration of coal mine spoils

# Selected Publications

A complete list of publications is available from my ORCID (<https://orcid.org/0000-0003-2252-7623>)

## Journal articles

```{r}
# bibliography_entries("mypubs.bib") |> 
#     filter(type == "article-journal") |>
#     arrange(desc(issued))
```

- Kumar, P., **Kumar, A.**, Patil, M., Hussain, S., & Singh, A. N. (2024). Factors influencing tree biomass and carbon stock in the Western Himalayas, India. *Frontiers in Forests and Global Change, 6*, 1328694. <https://doi.org/10.3389/ffgc.2023.1328694>

- Singh, A. N., & **Kumar, A.** (2022). Comparative soil restoration potential of exotic and native woody plantations on coal mine spoil in a dry tropical environment of India: A case study. *Land Degradation & Development, 33*(12), 1971--1984. <https://doi.org/10.1002/ldr.4286>

- Singh, A. N., & **Kumar, A.** (2022). Ecological performances of exotic and native woody species on coal mine spoil in Indian dry tropical region. *Ecological Engineering, 174*, 106470. <https://doi.org/10.1016/j.ecoleng.2021.106470>

- **Kumar, A.**, Patil, M., Kumar, P., Kumar, M., & Singh, A. N. (2022). Plant ecology in Indian Siwalik range: A systematic map and its bibliometric analysis. *Tropical Ecology, 63*(3), 338--350. <https://doi.org/10.1007/s42965-022-00229-x>

- Patil, M., **Kumar, A.**, Kumar, P., Cheema, N. K., Kaur, R., Bhatti, R., & Singh, A. N. (2020). Comparative litter decomposability traits of selected native and exotic woody species from an urban environment of north-western Siwalik region, India. *Scientific Reports, 10*, 7888. <https://doi.org/10.1038/s41598-020-64576-2>


## Book chapters

- **Kumar, A.**, Patil, M., Kumar, P., & Singh, A. N. (2021). Phosphorus and litter decomposability traits in tropical forest ecosystems under changing environment: A synthesis. In R. K. Chaturvedi, R. Singh, & R. Bhadouria (Eds.), *Tropical dry forests: Emerging features and ecological perspectives* (pp. 311--336). Nova Science Publishers.

- **Kumar, A.**, Yadav, R., Patil, M., Kumar, P., Zhang, L., Kaur, A., Sharma, S., Hussain, S., Tokas, D., & Singh, A. N. (2020). Sustainable management of national parks and protected areas for conserving biodiversity in India. In L. Zhang (Ed.), *Advances in forest management under global change* (pp. 75--91). IntechOpen. <https://doi.org/10.5772/intechopen.92435>

# Skills and Qualifications

A full list of accomplishments and certificates is available from my webpage (<https://akumar.netlify.app>)

- `R` statistical environment and `RStudio`
- Data analysis and visualisation (`MuMIn`, `piecewiseSEM`, `stats`, `tidyverse`)
- Mapping and spatial data analysis (`ggplot2`, `sf`, `terra`, `tmap`)
- Species distribution modelling (`biomod2`, `MaxEnt`)
- Systematic reviews and meta-analysis (`bibliometrix`, `metafor`, `revtools`)
- Reproducible research (`git`, `quarto`, `rmarkdown`, `knitr`)
- Microsoft office (`Excel`, `PowerPoint`, `Word`)
- Reference management (`Mendeley`)
- Scientific and technical writing

# Professional Memberships

```{r}

tribble(
  ~Membership, ~Year, ~Institution, ~Where,
  "Student member (#521)", "2023-Present", "Society of Open, Reliable and Transparent Ecology and Evolutionary Biology (SORTEE)", "Keizer, Oregon, US",
  "Life member (#L-1599)", "2022-Present", "Punjab Academy of Sciences", "Patiala, Punjab, IN",
  "Student member (#67527876)", "2021-2025", "International Association for Vegetation Science (IAVS)", "Bethesda, Maryland, US"
) |> 
  detailed_entries(Membership, Year, Institution, Where)
```

# References

```{r}
# what, when, where, with, why
tribble(
  ~Referee, ~Address, ~Relation, ~Email,
  "Dr. Anand Narain Singh", "Panjab University, Chandigarh, IN", "PhD Supervisor", "dranand1212@gmail.com",
  "Prof. Amrik Singh Ahluwalia", "Eternal University, Sirmaur, IN", "MSc Supervisor", "amrik.s511@gmail.com",
  "Dr. Balkar Singh", "Arya PG College, Panipat, IN", "BSc Lecturer", "balkararya@gmail.com"
) |>
  detailed_entries(what = Relation, where = Address, with = Referee, when = Email)
```

